Global code motion of congruent instructions

### 2017-11-08 cgo/paper.tex
chi details

A CHI keeps information about values flowing out of a basic block.  It is
similar to $\phi$ \cite{ssa} but in the inverse graph, and used for out-flowing
values on each edge. Although it may be suitable for other
analyses/optimizations, for \GCM{}, it is computed only for instructions with
multiple occurrences in the CFG. An illustration of CHI is shown in
Figure~\ref{fig:chi-intro}. The \GVN{} for both I1 and I2 is V, the CHI node
will save the instruction as well as the edge where the value is flowing to.
Where CHIArg1, CHIArg2 are of type CHIArg. The CHI for both I1 and I2 is
inserted in their PDF (A). For sinking we use the similar approach as
\cite{ssapre} to track values flowing into a basic block.

\begin{figure}
  \includegraphics[width=0.4\textwidth]{chi-example.pdf}
  \vspace*{-1.0cm}
\caption{I1, I2 are instructions in B, and C respectively, CHIArg1 = \{B, I1,
  V\} and CHIArg2 = \{C, I2, V\}, V is the value number for both I1, and I2}
\label{fig:chi-intro}
\end{figure}

When SSAPRE inserts $\phi$ (expression PHIs), it does so at places where
(potentially) multiple expressions may merge. If a $\phi$ has a missing entry, that
means the expression is partially available.  The concept may work for hoisting
expression to help factor out anticipability by working on inverted graph. If we
insert a CHI, to a basic block with multiple successors, as instructions are
hoisted upwards to the post-dominance frontier.  And we start walking the CFG to
figure out if any CHI has a missing entry, that means the expression is not
anticipable in the basic block having CHI.

>>>>

A CHI keeps information about values flowing out of a basic block.  It is
similar to $\phi$ \cite{ssa} but in the inverse graph, and used for out-flowing
values on each edge. Although it may be suitable for other
analyses/optimizations, for \GCM{}, it is computed only for instructions with
multiple occurrences in the CFG. An illustration of CHI is shown in
Figure~\ref{fig:chi-intro}. The \GVN{} for both I1 and I2 is V, the CHI node
will save the instruction as well as the edge where the value is flowing to.
Where $Arg1$, $Arg2$ are of type CHIArg. The CHI for both I1 and I2 is
inserted in their PDF (A). For sinking we use the similar approach as
\cite{ssapre} to track values flowing into a basic block.

\begin{figure}
  \includegraphics[width=0.4\textwidth]{chi-example.pdf}
  \vspace*{-1.0cm}
\caption{I1, I2 are instructions in B, and C respectively, Arg1 = \{B, I1,
  V\} and Arg2 = \{C, I2, V\}, V is the value number for both I1, and I2}
\label{fig:chi-intro}
\end{figure}

When SSAPRE inserts $\Phi$ (expression PHIs), it does so at places where
(potentially) multiple expressions may merge. If a $\Phi$ has a missing entry,
that means the expression is partially available.  We apply a similar concept
for hoisting expression by working on inverted graph. We insert a CHI, to a
basic block with multiple successors. Then for each instruction, we register its
flow of value to its corresponding CHI which can be found in its PDF. For
conciseness we only register values with multiple occurrences. Finally, we start
walking the CFG and remove the CHI if it has a missing entry, that means the
expression is not anticipable in the basic block having CHI. The completeness of
CHI is a necessary condition for hoistability but not sufficient. This is
because congruence of value numbers only implies that they compute same value if
the inputs are same. So for instructions like memory references, they may
compute different values should the memory get modified along the path. For that
we need to check for intersecting side-effects along the path.

\begin{figure}
  \includegraphics[width=0.4\textwidth]{chi-example2.pdf}
  \vspace*{-1.0cm}
\caption{I1, I2 are instructions in B, and C respectively, Arg1 = \{B, I1, V\}
  and Arg2 = \{C, I2, V\}, V is the value number for both I1, and I2. A has
  missing entry in CHI so values are not anticipable}
\label{fig:chi-intro2}
\end{figure}
